---
title: "Galaxy API R Client"
output:html_notebook
---
## The Galaxy container image

[README](https://github.com/bgruening/docker-galaxy-stable) contains detailed instructions
for controlling the container behavior.

In particular, [this section](https://github.com/bgruening/docker-galaxy-stable#Galaxys-config-settings)
decsribes default values of the admin account name, and the way to override it.
The default is:
```
GALAXY_CONFIG_ADMIN_USERS=admin@galaxy.org
```
and the password is `admin`.

You can override in this way any parameter from the [galaxy config file](https://github.com/galaxyproject/galaxy/blob/release_19.01/config/galaxy.yml.sample). 
This is not a container feature, this is a core Galaxy feature. When its config parser sees
an environment variable GALAXY_CONFIG_XXX, it will use its value to override the `galaxy.yaml` 
parameter `xxx`.

## Starting the container(s)

### Galaxy container

- In a terminal, create a directory that will be bind-mounted to keep Galaxy data alive across container restarts: `mkdir galaxy_storage`
- `docker create -v /export --name galaxy-store bgruening/galaxy-stable /bin/true`
   This is a big image - will take a while to pull.
- `docker run -d -p 8080:80 -p 8800:8800 --rm --name galaxy --privileged --volumes-from galaxy-store bgruening/galaxy-stable`
  We use `--privileged` to auto-mount reference data through CVMFS (works on Windows?)
- Access container in Chrome at `localhost:8080`.
- Go to User->Preferences->Manage API key. Click Create a new key. 
  Copy the key somewhere to use in your client R session.

### R Studio Container

*You can also use some other R, but we installed a couple of necessary packages into our container*

- `docker run -v /home/rstudio/work2 --name rstudio-work sliders/rstudio-bcbiornaseq:v2 bash -c "chown -R rstudio. /home/rstudio/work2"`
  We created a volume owned by the user `rstudio` from the container
- `docker run -d -p 8787:8787 -e PASSWORD=bcbio -e ROOT=TRUE --volumes-from rstudio-work  sliders/rstudio-bcbiornaseq:v2`
- Access the R Studio Server at `localhost:8787`
- Login with user `rstudio` and password `bcbio`
- Open Tools->Terminal, and `cd work2 && git clone https://github.com/andreyto/galaxy_r_api_workshop.git`
- `sudo apt-get install python-pip`
- `sudo pip install -U git+https://github.com/andreyto/parsec.git@at_fixes`
  Installs my branch of that package with some bug fixes

## Working in R

The API and bindings from different languages are described [here](https://galaxyproject.org/develop/api/)

### Define API key and URL
```{r}
GAL_URL="http://10.25.117.136:8080"
GAL_API_KEY="73c730a5ac10f80d363bcd46b4ca28a8"
```

At first, we will still run command from the terminal:

Run init and type your URL and key
```
parsec init
```
Download dataset by History Content API ID
```
parsec datasets download_dataset --file_path . 5969b1f7201f12ae
```
```{r}
data = readRDS("Galaxy1-[copd_brushings_rnaseq_bcb.rds].binary")
data
```
## Build Galaxy tool that works on our serialized BCBio SummarizedExperiment dataset

This is a very primitive prototype. We do not define a datatype and use default `binary` format for the `rds` file.

The primary objective is to demonstrate running a simple `Rscript` tool using our custom
Docker container to provide the dependencies.

We bind-mount several Galaxy config files and the tool itself into the container,  
and override the necessary locations through GALAXY_CONFIG_xxx environment variables passed
into the container. We do it in a fairly simplistic way just for demonstration.

In any case, the real-life tool code development is much better done through `Planemo` and
deployment through a local toolshed.

**Note**: To provide absolute host paths for bind-mounting, we use Unix shell command `pwd`,
which is not avalibale on Windows. You will have to edit the corresponding paths in the `docker run`
command if you are on Windows. Better still, active Linux subsystem in Windows 10 and use Bash shell.

### In a terminal on your *host*:
- `git clone https://github.com/andreyto/galaxy_r_api_workshop.git`
- `docker stop galaxy`
- ```
  docker run -d --rm -p 8080:80 -p 8800:8800 --name galaxy --privileged --volumes-from galaxy-store \
  -v `pwd`/galaxy_r_api_workshop:/extra \
  -e GALAXY_CONFIG_TOOL_CONFIG_FILE="/extra/config/tool_conf.xml,config/tool_conf.xml.sample,config/shed_tool_conf.xml" \
  -e GALAXY_CONFIG_JOB_CONFIG_FILE="/extra/config/job_conf.xml" \
  bgruening/galaxy-stable \
  bash -c "supervisorctl stop docker; service docker stop; service docker start; startup"
  ```

**Notes**: 
- When Docker inside the container pulls images, they do not get saved between container
  restarts. Expect to wait each time after a restart for the full pull when you run the tool
  for the first time.
- Sometimes, for reasons unclear, you can get an error status from the Galaxy tool, with the message
  saying that docker is not available (inside the container). This happens after the tool job was in
  a running (yellow) state for a while, and the Docker was pulling the container for the dependency.
  Apparently, Docker daemon dies inside the container.
  Open interactie shell inside the contaioner using `docker exec -it galaxy bash` and start Docker there
  again with `service docker stop; service docker start`. Then, rerun the Galaxy tool.
